---
title: "Homework week 2"
author: "Antoine Godin"
date: "October 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Using pdfetch

To show how to use pdfetch, I'm going to use a simple exercise: how to compute the income to wealth ratio of households for many countries at once. First you need to load the library

```{r, echo=T,eval=T}
library(pdfetch)
```

In this specific example, I will need to get data from different databases: flow data from the non-financial transaction for consumption and two measures of disposable income (gross and net), stock data from the financial balance sheet: total financial assets and stock data from the non-financial balance sheet: housing wealth.

We will first get the data from the non-financial transactions
```{r, echo=T,eval=T}
# Selecting the flows for nasa_nf_tr
names<-c("B6G","B6N","P3")
#getting the data in a raw format
NF_TR_raw = pdfetch_EUROSTAT("nasa_10_nf_tr", UNIT="CP_MNAC", NA_ITEM=names, from="1994-12-31", to="2016-12-31",
														 SECTOR=c("S14_S15"), DIRECT = "PAID")
#Transforming the data into a data.frame (easy to access and to view)
NF_TR<-as.data.frame(NF_TR_raw)
```

Then we look for non-financial balance sheets
```{r, echo=T,eval=T}
#selection of the flows for nama_10_nfa_bs (N11N - dwellings - net)
names<-c("N11N")
NFA_BS_raw = pdfetch_EUROSTAT("nama_10_nfa_bs", UNIT="CP_MNAC",ASSET10=names, from="1994-12-31", to="2016-12-31", SECTOR=c("S14_S15"))
NFA_BS<-as.data.frame(NFA_BS_raw)
```

And finally the financial balance sheet
```{r, echo=T,eval=T}
#selection of the flows for nasa_10_f_bs (BF90 - net financial assets)
names<-c("BF90")
F_BS_raw = pdfetch_EUROSTAT("nasa_10_f_bs", UNIT="MIO_NAC",CO_NCO="CO",NA_ITEM=names, from="1994-12-31", to="2016-12-31",
														SECTOR=c("S14_S15"))
F_BS<-as.data.frame(F_BS_raw)
```

If you look at the different data frame generated, you will see weird column names. This is due to the way pdfetch works, he concatenates all the information regarding a timeseries (all the meta-data: geo, na_item, unit, etc...) into one name, separated by dots. This is not realy readable, so I created a little sub-routine to clean those column name and only keep the indicator and the country code. Because each data frame has different names, I had to update each routine for the data frame. While I'm doing this, I'm also creating a new dataset called HHdata which concatenates all the information coming from the 3 datasets. The only issue with this is the length of the varisous timeseries, so I had to truncatenate some of these. The last thing I'm doing is to collect all the country code I can get, to be used afterwards.
```{r, echo=T,eval=T}
# Automatic procedure to remove the non-interesting bit of the colnames
HHdata<-c()
newcoln<-c()
countries<-c()

# For the the NF_TR data.frame
coln<-colnames(NF_TR)
for(i in 1:length(coln)){
	name<-coln[i]
	tname<-strsplit(name,"\\.")[[1]]
	newname<-paste(tname[c(4,6)],collapse=".")
	#Adding the current country code to the list of country codes
	countries<-c(countries,tname[6])
	# If the column contains only NA, remove it from the dataset
	if(sum(is.na(NF_TR[,i]))!=length(NF_TR[,i])){
		newcoln<-c(newcoln,newname) 
		#creating the HHdata dataset by column binding the exisiting HHdata with the new column
		HHdata<-cbind(HHdata,NF_TR[,i])
	}
}

# For the the NFA_BS data.frame
coln<-colnames(NFA_BS)
for(i in 1:length(coln)){
	name<-coln[i]
	tname<-strsplit(name,"\\.")[[1]]
	newname<-paste(tname[c(4,5)],collapse=".")
	#Adding the current country code to the list of country codes
	countries<-c(countries,tname[5])
	# If the column contains only NA, remove it from the dataset
	if(sum(is.na(NFA_BS[,i]))!=length(NFA_BS[,i])){
		newcoln<-c(newcoln,newname) 
		#creating the HHdata dataset by column binding the exisiting HHdata with the new column
		HHdata<-cbind(HHdata,NFA_BS[,i])
	}
}

# For the the F_BS data.frame
coln<-colnames(F_BS)
for(i in 1:length(coln)){
	name<-coln[i]
	tname<-strsplit(name,"\\.")[[1]]
	newname<-paste(tname[c(6,7)],collapse=".")
	#Adding the current country code to the list of country codes
	countries<-c(countries,tname[7])
	# If the column contains only NA, remove it from the dataset
	if(sum(is.na(F_BS[,i]))!=length(F_BS[,i])){
		newcoln<-c(newcoln,newname) 
		#creating the HHdata dataset by column binding the exisiting HHdata with the new column
		HHdata<-cbind(HHdata,F_BS[,i])
	}
}
```

Now, I will transform the HHdata into a data.frame for ease of use and give nice names to column and rows.
```{r, echo=T,eval=T}
HHdata<-as.data.frame(HHdata)
colnames(HHdata)<-newcoln
rownames(HHdata)<-rownames(F_BS)
```

Now I want to plot the data I'm interested in for all the countries for which I can. Remember the `countries` vector containing all the country codes? I am going to extract all the codes by using the `unique` function which gives me each code only once. Then I will apply the plot procedure for each country.
```{r, echo=T, eval=T, fig.width=14}
# getting the unique country codes
countries<-unique(countries)

#For each country
for(country in countries){
	#Get all the column in which the country code exists
	indices<-grep(country,newcoln)
	#Plotting the data only if I have the 5 timeseries I need
	if(length(indices)==5){
		#This will create a jpg with the country code as name
		#jpeg(filename=paste(country,".jpg",sep=""))
		#Dividing the plot into 4 sub-plots (2 by 2)
		#par(mfrow = c(2,2))
		#Plot 1 gross disposable income over total wealth
		timeseries<-HHdata[,paste("B6G",country,sep=".")]/
			(HHdata[,paste("N11N",country,sep=".")]+HHdata[,paste("BF90",country,sep=".")])
		plot(as.numeric(substr(rownames(HHdata),1,4)),timeseries,type="l",main=paste("B6G/Wealth",country),xlab = "",ylab="")
		grid()
		#Plot 2 net disposable income over total wealth
		timeseries<-HHdata[,paste("B6N",country,sep=".")]/
			(HHdata[,paste("N11N",country,sep=".")]+HHdata[,paste("BF90",country,sep=".")])
		plot(as.numeric(substr(rownames(HHdata),1,4)),timeseries,type="l",main=paste("B6N/Wealth",country),xlab = "",ylab="")
		grid()
		#Plot 3 total consumption over total wealth
		timeseries<-HHdata[,paste("P3",country,sep=".")]/
			(HHdata[,paste("N11N",country,sep=".")]+HHdata[,paste("BF90",country,sep=".")])
		plot(as.numeric(substr(rownames(HHdata),1,4)),timeseries,type="l",main=paste("P3/Wealth",country),xlab = "",ylab="")
		grid()
		#Plot 4 total consumption over net disposable income
		timeseries<-HHdata[,paste("P3",country,sep=".")]/
			(HHdata[,paste("B6N",country,sep=".")])
		plot(as.numeric(substr(rownames(HHdata),1,4)),timeseries,type="l",main=paste("P3/B6N",country),xlab = "",ylab="")
		grid()
		#This finished the jpeg creation
		#dev.off()
	}
}
```

After running the code (all contained into the `Consumption function.R` file), you should have 17 jpegs created in your working directory.

## Godley Stock-Flow norms

Your homework: I want you to compare Godley's stock flow norms (see his paper) for all the european countries for which there is enough data. Generate jpg and we will comment them in the next lecture.